---
title: "Comparison of Political Affiliation & Identity as an Environmentalist"
author: "Kaley Dodson"
date: "12/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(kableExtra)
library(here)
library(janitor)
library(broom)

# Read in data
conservation_survey <- read.csv(here("data", "conservation_survey.csv")) %>% 
  clean_names()
```

### 1. Introduction

This mini-report uses data from a survey conducted by Chelsea Batavia and Andrews Forest LTER to explore "Influences on charitable giving for conservation" to describe the association between respondents’ political affiliation and if they consider themselves an environmentalist. Survey responses to the prompts "I consider myself an environmentalist" and "How do you self-identify politically?" are counted, visualized in a table, and compared by chi-square test using a significance level ($\alpha$) of 0.05.

```{r, include = FALSE}
# subset of survey questions
subset_conservation_survey <- conservation_survey %>% 
  select(env_id, pol_or) %>% # retains variables for environmentalist identity and political orientation questions only
  mutate(pol_party = case_when( # adds new column which contains:
    pol_or %in% c(1, 2, 3) ~ "Conservative", # "conservative" if response is 1-3
    pol_or %in% c(5, 6, 7) ~ "Liberal", # "liberal" if response is 5-7
  )) 
```

### 2. Exploratory Findings & Summary

**Table 1.** Relation between identity as an environmentalist and political leaning. Values indicated are percentages with counts shown in parentheses. Data: Andrews Forest LTER Site and C. Batavia. (2019)

```{r, echo = FALSE}
# table 1 code

# create counts table object
survey_counts <- subset_conservation_survey %>% 
  tabyl(pol_party, env_id) 

# add proportions to counts table
survey_counts_prop <- survey_counts %>% 
  adorn_percentages() %>% # converts to percent
  adorn_pct_formatting(digits = 2) %>% # rounds to two digits
  adorn_ns() %>% # includes actual counts
  drop_na() # removes na values

# counts table formatting
survey_counts_prop %>% 
  kable(col.names = c(" ",
                      "Yes",
                      "No",
                      "Unsure")) %>% 
  kable_styling(full_width = FALSE)
```

```{r, include = FALSE}
# convert entries in table columns to stored rownames
survey_ct <- survey_counts %>% 
  drop_na %>% 
  column_to_rownames(var = "pol_party") 

# chi-square test
survey_x2 <- chisq.test(survey_ct)
survey_x2
# tidy model results for in-line reference
survey_tidy <- tidy(survey_x2)
```

There is a significant association between survey respondents’ political affiliation and whether they consider themselves an environmentalist ($\chi$^2^(`r survey_tidy$parameter`) = `r round(survey_tidy$statistic,2)`, *p* = `r format(survey_tidy$p.value, scientific = TRUE, digits = 3)`). The majority of those who leaned liberal also self-identified as an environmentalist (`r survey_counts_prop$"1"[2]`) while the majority of conservative-leaning respondents specifically did not (`r survey_counts_prop$"2"[1]`).

### Citation

Andrews Forest LTER Site and C. Batavia. 2019. Influences on charitable giving for conservation: Online survey data of 1,331 respondents across the US, August 2017 ver 3. Environmental Data Initiative. https://doi.org/10.6073/pasta/926e6270e324a1322a900da14d38b96c

